Modeling Human Interactions in Collaborative Interactive Evolutionary Computation

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Back to 8 pages, complete with three case studies and distribution fit

  \subsection{EvoSpace-Redis}
EvoSpace is a population store \cite{Evospace}  for the development of 
evolutionary algorithms that are intended to run on a cloud computing model. 
The population is decoupled from any particular evolutionary algorithm. 
Candidate solutions are stored as of objects in a population, and they can be withdrawn, 
processed and replaced using a specified set of methods \cite{GValdez2015}. The population
is stored in-memory, using the Redis key-value database. Redis was chosen over a relational 
database management system, because it provides a hash based implementation of sets and 
queues which are natural data structures for the EvoSpace model. Basicaly a sample of 
candidate solutions are retieved from the server, evaluated, and then sent back. 
The same operations are used to Evolve the population. In EvoSpace individuals replaced 
in the population are stored indefinitly, to permit users to store permalinks to them.
Implementation details are presented in \cite{garcia2013evospace}.

\subsection{Graph-Neo4J}
 Collaborative IEC (C-IEC) systems need to store highly connected data, as it is common 
 in current applications like social networks. In order to deal with large datasets of connected 
 data found in these systems, graph databases \cite{angles2012comparison} have been proposed 
 as an alternative to relational databases which have performance limitations when dealing with 
 highly connected data \cite{holzschuher2013performance}.
 A graph is proposed for modeling the social network of users, their interaction with 
 candidate solutions, and the relationships between them in the population.
 The graph database system used in the implementation is Neo4J as it is
 a scalable solution \cite{miller2013graph,holzschuher2013performance}, and well 
 supported and documented in PaaS infrastructures like Heroku. The Cypher query 
 language it used to retrieve views from the graph.
 An exaplmple query is shown in figure \ref{fig:cypher} where the relations 
 between users and solutions are presented.

@@

\subsection{Relational Objects-PostgreSQL}
  The PostgreSQL relational database system is also employed because user 
  sessions and authentication,
  as well as dynamic web pages are handled directly by the Django web
  framework \cite{garcia2013evospace}. % these implementations details
                                % could go if needed.

\subsection{Interfaces}
The appropiate human interface will depend on the application domain, 
the current framework implemantation employs a web based application. Developers 
can use different templates, to present one or more phenotypes 
at a time, and the type of rating system: like or one to five stars. 
Another implementations do not need a graphical interface at all, for example the
EvoDraw instalation or the XYZ Project \cite{de2014artists}.

\subsection{Evolution}
The evolutionary algorithm is decoupled from the data model, the algorithm 
could be implemented as an EvoWorker \cite{garcia2013evospace}, use an external library, 
or even assigned to humans. An additional tool was implmented for a human based evolution, 
where volunteers select the parents of the next generation and upload the 
new individuals \cite{de2014artists}.

@@

\begin{table}
  \small
  \caption{ Parameters for experiments.  }
  \label{tab:params} 
  \centering
  \small
  \begin{tabular}{l  c   }
    \hline\noalign{\smallskip}
     Parameter & Value \\
    \noalign{\smallskip}\hline\noalign{\smallskip}
    Initial Population Size   & 80 \\ \hline
    Sample Size & 1 \\ \hline
    Step Size & 8 Samples \\ \hline
    Mutation &  \\ \hline
    Selection & Tournament \\ \hline
    Tournament Size &  6 \\ \hline
  \end{tabular}
\end{table}

@@

In orther to assess if C-IEC follow the same pattern, even when
using gamification techniques, participation was fittted to a 
Weibull distribution, results in figures \ref{fig:w1}--\ref{fig:w3}  show that 
the same pattern appears in this case study.

\begin{figure}[!t]
    \centering
        \includegraphics[width=3in]{img/weibull_1.png}
    \caption{Experiment 1. Weibull Fit}
    \label{fig:w1}
\end{figure}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3in]{img/weibull_2.png}
    \caption{Experiment 2. Weibull Fit}
    \label{fig:w2}
\end{figure}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3in]{img/weibull_3.png}
    \caption{Experiment 3. Weibull Fit}
    \label{fig:w3}
\end{figure}

@@

>>>>

\begin{itemize}
\item {\bf EvoSpace-Redis} 
EvoSpace is a population store \cite{Evospace}  for the development of 
evolutionary algorithms that are intended to run on a cloud computing model. 
The population is decoupled from any particular evolutionary algorithm. 
Candidate solutions are stored as of objects in a population, and they can be withdrawn, 
processed and replaced using a specified set of methods \cite{GValdez2015}. The population
is stored in-memory, using the Redis key-value database. Redis was chosen over a relational 
database management system, because it provides a hash based implementation of sets and 
queues which are natural data structures for the EvoSpace model. Basicaly a sample of 
candidate solutions are retieved from the server, evaluated, and then sent back. 
The same operations are used to Evolve the population. In EvoSpace individuals replaced 
in the population are stored indefinitly, to permit users to store permalinks to them.
Implementation details are presented in \cite{garcia2013evospace}.

\item {\bf Graph-Neo4J} 
 Collaborative IEC (C-IEC) systems need to store highly connected data, as it is common 
 in current applications like social networks. In order to deal with large datasets of connected 
 data found in these systems, graph databases \cite{angles2012comparison} have been proposed 
 as an alternative to relational databases which have performance limitations when dealing with 
 highly connected data \cite{holzschuher2013performance}.
 A graph is proposed for modeling the social network of users, their interaction with 
 candidate solutions, and the relationships between them in the population.
 The graph database system used in the implementation is Neo4J as it is
 a scalable solution \cite{miller2013graph,holzschuher2013performance}, and well 
 supported and documented in PaaS infrastructures like Heroku. The Cypher query 
 language it used to retrieve views from the graph.
 An exaplmple query is shown in figure \ref{fig:cypher} where the relations 
 between users and solutions are presented.

@@

\item {\bf Relational Objects-PostgreSQL}
  The PostgreSQL relational database system is also employed because user 
  sessions and authentication,
  as well as dynamic web pages are handled directly by the Django web
  framework \cite{garcia2013evospace}. % these implementations details
                                % could go if needed.

\item {\bf Interfaces}
The appropiate human interface will depend on the application domain, 
the current framework implemantation employs a web based application. Developers 
can use different templates, to present one or more phenotypes 
at a time, and the type of rating system: like or one to five stars. 
Another implementations do not need a graphical interface at all, for example the
EvoDraw instalation or the XYZ Project \cite{de2014artists}.

\item {\bf Evolution}    
The evolutionary algorithm is decoupled from the data model, the algorithm 
could be implemented as an EvoWorker \cite{garcia2013evospace}, use an external library, 
or even assigned to humans. An additional tool was implmented for a human based evolution, 
where volunteers select the parents of the next generation and upload the 
new individuals \cite{de2014artists}.  
\end{itemize}

@@

\begin{table}
  \small
  \caption{ Parameters for experiments.  }
  \label{tab:params} 
  \centering
  \small
  \begin{tabular}{l  c}
    \hline\noalign{\smallskip}
     Parameter & Value \\
    \noalign{\smallskip}\hline\noalign{\smallskip}
    Initial Population Size   & 80 \\ \hline
    Sample Size & 1 \\ \hline
    Step Size & 8 Samples \\ \hline
    Mutation &  \\ \hline
    Selection & Tournament \\ \hline
    Tournament Size &  6 \\ \hline
  \end{tabular}
\end{table}

@@

In orther to assess if C-IEC follow the same pattern, even when
using gamification techniques, participation was fittted to a 
Weibull distribution, results in figure \ref{fig:w2} and figure \ref{fig:w3}
show that the same pattern appears in this case study.

\begin{figure}[!t]
    \centering
        \includegraphics[width=3in]{img/weibull_2.png}
    \caption{Experiment 2. Weibull Fit}
    \label{fig:w2}
\end{figure}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3in]{img/weibull_3.png}
    \caption{Experiment 3. Weibull Fit}
    \label{fig:w3}
\end{figure}

@@

Collaborative and volunteer based IEC systems involve the dynamic 
interaction of many entities and artifacts. Employing human centered approach 
can benefit researchers to understand and visualize this kind of systems better.
In this paper a human-centered software framework was proposed, and was validated
through the implementaion and refinement of three C-IEC applications.
This framework enabled the implementation of a gamification technique 
to improve engagement in a case study. Providing a common arena where users are aware of the
activities of other users in the social vicinity. Following a hybrid data model approach
gives developers and project owners options when innovating with other means
of interaction or new applications.

One of the interesting future lines of work would be to look a bit
more closely at the behavior of users as they are rating artifacts 
in the web system. These initial experiments hint at a possible power law, which might indicate that
the IEC system could be self-organizing, a process that would allow it
to reach a critical state, as has been found in software repositories,
for instance \cite{Merelo2016:repomining}. 
The dynamics of this kind of system are fundamentally different, and our future research will
include exploring these aspects of the system.

Another line of work would be to study the negative effects of using 
gamification techniques to improve engagement, like cheating or competition.
Finally, the refinement of the proposed Human-Centered framework will need
more case studies and further multi-disciplinary research.
